---
title: "Compare Cori et al., Wallinga & Teunis and Bettencourt & Ribeiro methods"
output: html_notebook
---


```{r}
## Preamble
rm(list = ls())
library(ggplot2)
library(dplyr)
library(tidyr)
library(readr)
library(cowplot)
library(EpiEstim)
theme_set(theme_bw())
knitr::opts_chunk$set(message = FALSE)
parlist <- read_rds('true_pars.rds')
source('../code/util.R')
if(!dir.exists('figs')){dir.create('figs')}
```

## Extract SEIR simulation and plot.

- Incidence here is observed at time of infection (S -> E).
- SEIR simulation is deterministic

```{r}
get_sim_df() %>%
    filter(time < 300) %>%
ggplot() +
  geom_line(aes(x = time, y = incidence))+
  geom_vline(aes(xintercept = parlist$intervention_time_1), lty = 2)+ ## Dahsed line where Rt starts to decrease
    geom_vline(aes(xintercept = parlist$intervention_time_2), lty = 2)+ ## Dahsed line where Rt starts to decrease
  ggtitle('Epidemic curve') -> inc

get_sim_df() %>% 
  filter(time < 300) %>%
  ggplot()+
  geom_line(aes(x = time, y = true_rt)) +
  geom_hline(aes(yintercept = 1), lty = 2)+
  ylab(expression(paste(R[t])))+
  ggtitle(expression(paste('Underlying ', R[t], ' values'))) -> R0

cowplot::plot_grid(R0, inc, align = 'hv', nrow = 2)
ggsave('figs/SEIR_sim.png', width = 5, height = 5, units = 'in', dpi = 300)
```


# Fig. 2 - estimate Rt in real time using three different methods, assuming perfect observation at the moment of infection.

## Truncate the synthetic data to a maximum time of observation

```{r}
mt <- 150
rtdf <- get_sim_df() %>% filter(time <= mt)
```


## Use the method of Cori et al. and of Wallinga & Teunis from EpiEstim to estimate Rt.
```{r}
wdw <- 1 ## Set smoothing window = 1d
reset = TRUE ## If FALSE, read in cached file

if(!file.exists('cori_wt_fits_rt.rds')|reset){ ## If generating new estimates...
  plot_df_rt <- rtdf %>%
    merge(get_cori(rtdf, icol_name = 'incidence', window = wdw), by = 'time') %>%     ## Estimate using Cori et al. Function get_cori is defined in `util.R`
    merge(get_WT(rtdf, icol_name = 'incidence', window = wdw), by = 'time') %>%      ## Estimate using WT. Function get_WT is defined in `util.R` -- Note, this takes several minutes to generate CIs.
    merge(get_WT(rtdf, icol_name = 'obs_cases', outcol_name = 'WT_symptom_onset', window = wdw), by = 'time') %>%      ## Estimate using WT at E->I transition
    select(time, true_rt, incidence, obs_cases, contains('Cori'), contains('WT')) 
  
  write_rds(plot_df_rt, path = 'cori_wt_fits_rt.rds')
}else{ ## If reading saved file...
  plot_df_rt <- readRDS('cori_wt_fits_rt.rds')
}
```

## Get B & R fits, estimated using a model implemented in rstan.

```{r message = FALSE}
## wrapper function get_BR is defined in `util.R`.
## stan model is defined in rtlive.stan, and wrappers are defined in rtlive.R
BR_rt <- get_BR(rtdf, 'BR_real_time.rds', parlist, reset = TRUE)

## Merge with WT and Cori estimates from above
plot_rt <- merge(
  plot_df_rt,
  parse_fits(BR_rt, max.time = mt),
  by = 'time', all = TRUE
) %>%
  pivot_longer(contains('.'), names_to = c("Fit_method", '.value'), names_sep = '\\.') ## Reformat for plotting
```

## Calculate the case reproductive number
```{r}
source('../code/caseR.R')
beta_t <- with(get_sim_df(),
               approxfun(x = time, y = true_r0/parlist$t_I))

Rt_output <- with(parlist,
                  integrate_Rt(beta_t, 
                          sigma = 1/t_E, 
                          gamma = 1/t_I, 
                          N = N, 
                          T = mt, 
                          E0 = E_init, 
                          I0 = I_init, 
                          dt_int = .001)
) 
```

Merge Rc with the data frame for plotting
```{r}
## Add to plotting data frame.
plot_rt <- merge(Rt_output, plot_rt, by = 'time', all.y = TRUE)
```


## Plot Figure 2
```{r}
ymin = 0; ymax =2.5 ## Set y limits (truncate CIs if they exceed these limits)
plot_rt %>%
  filter(!is.na(mean) & time < 250) %>%
  mutate(`975` = ifelse(`975`>ymax, ymax, `975`),
         `025` = ifelse(`025`<ymin, ymin, `025`)) %>%
  mutate(facets = factor(Fit_method, levels = c('BR','Cori','WT_symptom_onset','WT'), labels = c('Bettencourt & Ribeiro', 'Cori et al.', 'Wallinga & Teunis', 'Wallinga & Teunis'))) %>%
  ggplot() +
  geom_line(aes(x = time, y = true_rt), lwd = 1)+
  geom_line(aes(x = time, y = R_case), lty = 2) +
  geom_line(aes(x = time, y = mean, color = Fit_method), alpha = .7) +
  geom_ribbon(aes(x = time, ymin = `025`, ymax = `975`, fill = Fit_method), alpha = 0.3)+
  xlab('time') +
  ylab(expression(paste(R[t]))) +
  xlab('Time')+
  ylim(c(ymin, ymax))+
  facet_wrap(.~facets) +
  xlim(c(0, mt))+
  theme(legend.position = 'none')+
  labs(fill = 'Estimation method', color = "Estimation method")
  gg_png(ww=6, hh=2.5, 'figs/compare_estimators.png')
```

## Make inset of time series
```{r}
plot_rt %>%
  ggplot()+
  geom_line(aes(x = time, y = incidence))+
  geom_line(aes(x = time, y = obs_cases), color = 'gray')+
  ylab(NULL)+
  theme_classic()
gg_png(1.6, 1.5, 'figs/compare_estimators_inset.png')
```
