---
title: "Analyze synthetic data and generate plots"
output: html_notebook
---

## This script makes the following figures:

* Fig. 1 (Compare 3 methods under ideal conditions)
* Fig. 4 (Effect of window size)
* Fig. 5 (Effect of mean serial interval)

```{r}
## Preamble
rm(list = ls())
library(ggplot2)
library(dplyr)
library(tidyr)
library(readr)
library(cowplot)
library(EpiEstim)
theme_set(theme_bw())
knitr::opts_chunk$set(message = FALSE)
parlist <- read_rds('true_pars.rds')
source('../code/util.R')
if(!dir.exists('figs')){dir.create('figs')}
```

## Extract SEIR simulation and plot.

- Incidence here is observed at time of infection (S -> E).
- SEIR simulation is deterministic

```{r}
get_sim_df() %>%
    filter(time < 300) %>%
ggplot() +
  geom_line(aes(x = time, y = incidence))+
  geom_vline(aes(xintercept = parlist$intervention_time_1), lty = 2)+ ## Dahsed line where Rt starts to decrease
    geom_vline(aes(xintercept = parlist$intervention_time_2), lty = 2)+ ## Dahsed line where Rt starts to decrease
  ggtitle('Epidemic curve') -> inc

get_sim_df() %>% 
  filter(time < 300) %>%
  ggplot()+
  geom_line(aes(x = time, y = true_rt)) +
  geom_hline(aes(yintercept = 1), lty = 2)+
  ylab(expression(paste(R[t])))+
  ggtitle(expression(paste('Underlying ', R[t], ' values'))) -> R0

plot_grid(R0, inc, align = 'hv', nrow = 2)
ggsave('figs/SEIR_sim.png', width = 5, height = 5, units = 'in', dpi = 300)
```


# Fig. 2 - estimate Rt in real time using three different methods, assuming perfect observation at the moment of infection.

## Truncate the synthetic data to a maximum time of observation

```{r}
mt <- 150
rtdf <- get_sim_df() %>% filter(time <= mt)
```


## Use the method of Cori et al. and of Wallinga & Teunis from EpiEstim to estimate Rt.
```{r}
wdw <- 1 ## Set smoothing window = 1d
reset = FALSE ## If FALSE, read in cached file

if(!file.exists('cori_wt_fits_rt.rds')|reset){ ## If generating new estimates...
  plot_df_rt <- rtdf %>%
    merge(get_cori(rtdf, icol_name = 'incidence', window = wdw), by = 'time') %>%     ## Estimate using Cori et al. Function get_cori is defined in `util.R`
    merge(get_WT(rtdf, icol_name = 'incidence', window = wdw), by = 'time') %>%      ## Estimate using WT. Function get_WT is defined in `util.R` -- Note, this takes several minutes to generate CIs.
    merge(get_WT(rtdf, icol_name = 'obs_cases', outcol_name = 'WT_symptom_onset', window = wdw), by = 'time') %>%      ## Estimate using WT at E->I transition
    select(time, true_rt, incidence, obs_cases, contains('Cori'), contains('WT')) 
  
  write_rds(plot_df_rt, path = 'cori_wt_fits_rt.rds')
}else{ ## If reading saved file...
  plot_df_rt <- readRDS('cori_wt_fits_rt.rds')
}
```

## Get B & R fits, estimated using a model implemented in rstan.

```{r message = FALSE}
## wrapper function get_BR is defined in `util.R`.
## stan model is defined in rtlive.stan, and wrappers are defined in rtlive.R
BR_rt <- get_BR(rtdf, 'BR_real_time.rds', parlist, reset = FALSE)

## Merge with WT and Cori estimates from above
plot_rt <- merge(
  plot_df_rt,
  parse_fits(BR_rt, max.time = mt),
  by = 'time', all = TRUE
) %>%
  pivot_longer(contains('.'), names_to = c("Fit_method", '.value'), names_sep = '\\.') ## Reformat for plotting
```

## Get the case reproductive number
```{r}
# reset = FALSE
# if(!file.exists('Rc.rds')|reset == TRUE){
# source('../code/simulation.R')
# Rc <- get_Rc_ode(arnaught = get_sim_df()$true_r0, 
#                  t_E = parlist$t_E, 
#                  t_I = parlist$t_I, 
#                  N = parlist$N, 
#                  E_init = parlist$E_init, 
#                  I_init = parlist$I_init, 
#                  n_t = parlist$n_t)
# write_rds(Rc, path = 'Rc.rds')
# }else{
#   Rc <- readRDS('Rc.rds')
# }

source('../code/caseR.R')
beta_t <- with(get_sim_df(),
               approxfun(x = time, y = true_r0/parlist$t_I))
# # Test
# ttt <- runif(n = 301, 0, 300)
# get_sim_df() %>%
#   ggplot()+
#   geom_line(aes(x = time, y = true_r0/parlist$t_I))+
#   geom_point(aes(x = ttt, y = beta_t(ttt)), color = 'red')

Rt_output <- with(parlist,
                  integrate_Rt(beta_t, 
                          sigma = 1/t_E, 
                          gamma = 1/t_I, 
                          N = N, 
                          T = mt, 
                          E0 = E_init, 
                          I0 = I_init, 
                          dt_int = .001)
) 
```

```{r}
## Add to plotting data frame.
plot_rt <- merge(Rt_output, plot_rt, by = 'time', all.y = TRUE)
```


## Plot Figure 2
```{r}
ymin = 0; ymax =2.5 ## Set y limits (truncate CIs if they exceed these limits)
plot_rt %>%
  filter(!is.na(mean) & time < 250) %>%
  mutate(`975` = ifelse(`975`>ymax, ymax, `975`),
         `025` = ifelse(`025`<ymin, ymin, `025`)) %>%
  mutate(facets = factor(Fit_method, levels = c('BR','Cori','WT_symptom_onset','WT'), labels = c('Bettencourt & Ribeiro', 'Cori et al.', 'Wallinga & Teunis', 'Wallinga & Teunis'))) %>%
  ggplot() +
  geom_line(aes(x = time, y = true_rt), lwd = 1)+
  geom_line(aes(x = time, y = mean, color = Fit_method)) +
  geom_line(aes(x = time, y = R_case), color = 'gray') +
  geom_ribbon(aes(x = time, ymin = `025`, ymax = `975`, fill = Fit_method), alpha = 0.3)+
  #geom_line(data = Rc, aes(x = time, y = R_case), color = 'gray')+
  xlab('time') +
  ylab(expression(paste(R[t]))) +
  xlab('Time')+
  ylim(c(ymin, ymax))+
  facet_wrap(.~facets) +
  xlim(c(0, mt))+
  theme(legend.position = 'none')+
  labs(fill = 'Estimation method', color = "Estimation method")
  gg_png(ww=6, hh=2.5, 'figs/Figure2.png')
```

## Make inset of time series
```{r}
plot_rt %>%
  ggplot()+
  geom_line(aes(x = time, y = incidence))+
  geom_line(aes(x = time, y = obs_cases), lty = 2)+
  ylab(NULL)+
  theme_classic()
gg_png(1.6, 1.5, 'figs/Fig2_inset.png')
```

# Make panels of Fig. 5

## Fig. 5A
## Use method of Cori et al. Compare smoothing windows. Assign R_t to end of window.

```{r}
cori_wrapper_wdw <- function(ww){
  get_cori(get_sim_df('stochastic') %>% filter(time < 150), 
           icol_name = 'incidence',
           out_name = sprintf('wdw_%.1f', ww),
           window = ww, 
           SI_mean = parlist$true_mean_SI)
}

wdw_df <- get_sim_df() %>%
  ## Merge with true SI fit
  merge(cori_wrapper_wdw(1) , by = 'time') %>%
 # merge(cori_wrapper_wdw(3), by = 'time') %>%
  merge(cori_wrapper_wdw(7), by = 'time') %>%
  merge(cori_wrapper_wdw(14), by = 'time')# %>%
  #merge(cori_wrapper_wdw(21), by = 'time') 

```



```{r}
ymax = 3
wdw_df %>%
  pivot_longer(contains('.'), names_to = c('window', '.value'), names_pattern = 'wdw_(\\d\\d?\\.\\d)\\.(\\w+)') %>%
  mutate(window = as.numeric(window),
         w_lab = sprintf('window = %.0f d', window)) %>%
  mutate(`975` = ifelse(`975`>ymax, ymax, `975`)) %>%
  filter(time < 150) %>%
  ggplot()+
  geom_line(aes(x = time, y = true_rt), lwd = 1)+
  geom_ribbon(aes(x = time, ymin = `025`, ymax = `975`, fill = as.factor(window)), alpha = .3) +
  geom_line(aes(x = time, y = mean, color = as.factor(window)), alpha = .7)+
  ylab(expression(paste(R[t])))+
  xlab('Time')+
  theme(legend.position = c(.75, .7))+
  labs(color = 'Window size (d)', fill = 'Window size (d)')
gg_png(4, 3.5, 'figs/Fig5A_window_tend.png')
```


## Make Fig. 4B
## Cori: compare smoothing windows, assing R_t to middle of window

```{r}
## Wraper to get cori estimates for a given window size, ww
cori_wrapper_wdw_mid <- function(ww){
  get_cori(get_sim_df(method = 'stochastic') %>% filter(time < 150), 
           icol_name = 'incidence', ## Which column to use as input data
           out_name = sprintf('wdw_%.1f', ww), ## Name of output column will be "wdw_SIZE"
           window = ww,
           SI_mean = parlist$true_mean_SI, 
           wend = FALSE) ## Assign to midpoint, not end of window
}

wdw_df_mid <- get_sim_df(method = 'stochastic') %>%
  ## Merge with true SI fit
  merge(cori_wrapper_wdw_mid(1) , by = 'time') %>%
  #merge(cori_wrapper_wdw_mid(3), by = 'time') %>%
  merge(cori_wrapper_wdw_mid(7), by = 'time') %>%
  merge(cori_wrapper_wdw_mid(14), by = 'time') #%>%
  #merge(cori_wrapper_wdw_mid(21), by = 'time') 

```



```{r}
ymax = 3
wdw_df_mid %>%
  pivot_longer(contains('.'), names_to = c('window', '.value'), names_pattern = 'wdw_(\\d\\d?\\.\\d)\\.(\\w+)') %>%
  mutate(window = as.numeric(window),
         w_lab = sprintf('window = %.0f d', window)) %>%
  mutate(`975` = ifelse(`975`>ymax, ymax, `975`)) %>%
  filter(time < 150) %>%
  ggplot()+
  geom_line(aes(x = time, y = true_rt), lwd = 1)+
  geom_ribbon(aes(x = time, ymin = `025`, ymax = `975`, fill = as.factor(window)), alpha = .3) +
  geom_line(aes(x = time, y = mean, color = as.factor(window)), alpha = .7)+
  ylab(expression(paste(R[t]))) +
  xlab('Time')+
  theme(legend.position = c(.75, .7))+
  labs(color = 'Window size (d)', fill = 'Window size (d)')
gg_png(4, 3.5, 'figs/Fig5B_window_tmid.png')
```

## Plot time series inset
```{r}
get_sim_df('stochastic') %>% filter(time < 150) %>%
  ggplot()+
  geom_line(aes(x = time, y = incidence))+
  ylab('new infections')+
  theme_classic()
gg_png(2.1,2, 'figs/Fig5_inset.png')
```


# Fig. 6

## Cori: compare serial intervals


```{r}
wdw <- 1 ## Set smoothing window = 1d
cori_wrapper_SI <- function(SI_offset){
  SI <- parlist$true_mean_SI+SI_offset
  get_cori(get_sim_df(), icol_name = 'incidence', out_name = sprintf('SI_%.1f', SI), window = wdw, SI_mean = SI)
}

SI_df <- get_sim_df('stochastic') %>%
  ## Merge with true SI fit
  merge( cori_wrapper_SI(0) , by = 'time') %>%
  merge(cori_wrapper_SI(-2), by = 'time') %>%
  merge(cori_wrapper_SI(-1), by = 'time') %>%
  merge(cori_wrapper_SI(1), by = 'time') %>%
  merge(cori_wrapper_SI(2), by = 'time') 

```

```{r}
ymax = 3
SI_df %>%
  pivot_longer(contains('.'), names_to = c('SI', '.value'), names_pattern = 'SI_(\\d\\d?\\.\\d)\\.(\\w+)') %>%
  mutate(SI = as.numeric(SI)) %>%
  mutate(`975` = ifelse(`975`>ymax, ymax, `975`)) %>%
  filter(time < 150) %>%
  mutate(SI = ifelse(SI == 8, '8 (true value)', SI)) %>%
  mutate(SI = factor(SI, levels = c('6', '7', '8 (true value)', '9', '10'))) %>%
  ggplot()+
  geom_line(aes(x = time, y = true_rt), lwd = 1)+
  geom_line(aes(x = time, y = mean, color = SI))+
  geom_ribbon(aes(x = time, ymin = `025`, ymax = `975`, fill =SI), alpha = .1) +
  ylab(expression(paste(R[t])))+
  labs(color = 'Generation interval\nmean (d)', fill = 'Generation interval\nmean (d)') +
  theme(legend.position = c(.75, .7))
gg_png(4, 3.5, 'figs/Fig6_gen_interval.png')
```



## Compare smoothed WT and Cori

```{r}
reset = TRUE ## If FALSE, read in cached file

if(!file.exists('smoothed.rds')|reset){ ## If generating new estimates...
  plot_smoothed <- rtdf %>%
    merge(get_cori(rtdf, icol_name = 'incidence', out_name = 'Cori_3', window = 3, wend = FALSE), by = 'time') %>%    ## Estimate using Cori et al. Function get_cori is defined in `util.R`
    merge(get_WT(rtdf, icol_name = 'incidence', outcol_name = 'WT_3', window = 3, wend = FALSE), by = 'time') %>%      ## Estimate using WT. Function get_WT is defined in `util.R` -- Note, this takes several minutes to generate CIs.
    merge(get_cori(rtdf, icol_name = 'incidence', out_name = 'Cori_7', window = 7, wend = FALSE), by = 'time') %>%     ## Estimate using Cori et al. Function get_cori is defined in `util.R`
    merge(get_WT(rtdf, icol_name = 'incidence', outcol_name = 'WT_7', window = 7, wend = FALSE), by = 'time') %>%      ## Estimate using WT. Function get_WT is defined in `util.R` -- Note, this takes several minutes to generate CIs.
    select(time, true_rt, incidence, contains('Cori'), contains('WT')) 
  
  write_rds(plot_smoothed, path = 'smoothed.rds')
}else{ ## If reading saved file...
  plot_smoothed <- readRDS('smoothed.rds')
}
```



```{r}
plot_smoothed %>%
  pivot_longer(-c(1:3), names_to = c('method', '.value'), names_sep = '\\.') %>%
  separate(method, into = c('method', 'window'), sep = '_') %>%
  filter(window == '7') %>%
  ggplot(aes(x = time)) +
  geom_line(aes(y = mean, color = method))+
    geom_ribbon(aes(x = time, ymin = `025`, ymax = `975`, fill = method), alpha = .5) +
  #geom_line(aes(y = true_rt))+
  #facet_wrap(.~window)+
  ylim(c(.5, 2.5))+
  ylab(expression(paste(R[t])))
gg_png(ww = 5.5, hh = 4, 'figs/Fig_compare_smooth.png')
```




```{r}
plot_smoothed %>%
  pivot_longer(-c(1:3), names_to = c('method', '.value'), names_sep = '\\.') %>%
  separate(method, into = c('method', 'window'), sep = '_') %>%
  mutate(stime = ifelse(method == 'Cori', time, time +8)) %>%
  ggplot(aes(x = stime)) +
  geom_line(aes(y = mean, color = method))+
  #geom_line(aes(y = true_rt))+
  facet_wrap(.~window)+
  ylim(c(.5, 2.5))
```