---
title: "Make Fig. 3, infer infections from observations"
output: html_notebook
---

Notebook by Katie Gostic 

Last updated: 5-14-2020

```{r}
rm(list = ls())
library(ggplot2)
library(dplyr)
library(tidyr)
library(readr)
library(cowplot)
library(EpiEstim)
theme_set(theme_bw())
knitr::opts_chunk$set(message = FALSE)
parlist <- read_rds('true_pars.rds')
source('../code/util.R')
```

## Visualize simulated data
```{r}
get_sim_df() %>%
ggplot() +
  geom_line(aes(x = time, y = incidence))+
  geom_vline(aes(xintercept = parlist$intervention_time_1), lty = 2)+ ## Dahsed line where Rt starts to decrease
    geom_vline(aes(xintercept = parlist$intervention_time_2), lty = 2)+ ## Dahsed line where Rt starts to decrease
  ggtitle('Daily incidence, SEIR simulation') -> inc

get_sim_df() %>% 
  ggplot()+
  geom_line(aes(x = time, y = true_r0)) +
  geom_hline(aes(yintercept = 1), lty = 2)+
  ylab('R0')+
  ggtitle('Underlying R0 values') -> R0

cowplot::plot_grid(R0, inc, align = 'hv', nrow = 2)
```
## Set the last date at which data observed
```{r}
mt <- 110
```

## Set the delay distribution

* Infection to case detection at case testing sites
* Infection to death

The assumed distributions are arbitrary for the purposes of testing, but for realism, I'm drawing on delay times reported in [Linton et al.][1], set the delay distribution from infection to case detection and to death. For case testing, assume the delay distribution is the incubation period (also from Linton et al), plus a 0-4 day delay from symptom onset to testing and reporting. \\
\\
```{r}
## Set the delay distribution to case testing
case_delay_dist <- function(nn){
  r_inc_dist <- function(n){rgamma(n, shape = 5.8, scale = 0.95)} # Incubation period (infection -> symptoms)
  r_sym_to_obs_dist <- function(n){runif(n, 0, 4)} # Additional delay from symptoms -> observation
  r_inc_dist(nn) + r_sym_to_obs_dist(nn)
}

## Use gamma parameterization from  Linton et al. 2020
##  https://www.mdpi.com/2077-0383/9/2/538
mean_d <- 15.0
var_d <- 6.9^2
## Set the delay distribution to death
death_delay_dist <- function(nn){
  r_inc_dist <- function(n){rgamma(n, shape = 5.8, scale = 0.95)} # Incubation period (infection -> symptoms)
  r_sym_to_obs_dist <- function(n){rgamma(n, shape = mean_d^2/var_d, rate = mean_d/var_d)} # Additional delay from symptoms -> observation
  r_inc_dist(nn) + r_sym_to_obs_dist(nn)
}
```


## Get synthetic data and truncate to observations up to the maximum time
```{r}
df.in <- get_sim_df() %>% filter(time <= mt)
```

## Forward step: Infer times of observation from SEIR times of infection

Implicitly we assume all infections in these data are later observed. This would not be the case in reality, but for illustration, we can interpret the underlying times of infection as the true times of infection for individuals who were later observed.

```{r}
source('../code/infer_times_of_infection_observation.R')

df.in %>%
  merge(
      ## Get times of case observation
    get_tObs_from_tInf(df.in$incidence, df.in$time, case_delay_dist, return_times = T),
    by = 'time', all = TRUE) %>% rename(obs_case = n) %>%
      ## Get observed times of death
  merge(
    get_tObs_from_tInf(df.in$incidence, df.in$time, death_delay_dist, return_times = T),
    by = 'time', all = TRUE) %>% rename(obs_deaths = n) %>%
  as.tbl() -> obs_df

```


## Backward imputation step: Get times of infection from times of observation

* Shift by the mean delay
* Or infer times of infection by sampling from the delay distribution

```{r fig.height}
backward_step_wrapper <- function(df.in, mt = mt){
  df.in %>%
    filter(time <= mt) %>%
  ## Impute times of observation by subtracting the mean or by drawing from the delay distribution
  mutate(shift_case = move_back_in_time(n_obs = obs_case, mean(case_delay_dist(1000))),
         shift_death = move_back_in_time(n_obs = obs_deaths, mean(death_delay_dist(1000))),
         sample_case = get_tInf_from_tObs(n_obs = obs_case, times = time, r_delay_dist = case_delay_dist, adjust_right_censor = TRUE),
         sample_case_unadjusted = get_tInf_from_tObs(n_obs = obs_case, times = time, r_delay_dist = case_delay_dist, adjust_right_censor = FALSE),
         sample_death = get_tInf_from_tObs(n_obs = obs_deaths, times = time, r_delay_dist = death_delay_dist),
         sample_death_unadjusted = get_tInf_from_tObs(n_obs = obs_deaths, times = time, r_delay_dist = death_delay_dist, adjust_right_censor = FALSE)) %>%
  select(time, true_rt, incidence, contains('obs'), contains('shift'), contains('sample')) 
}

inf_obs_df <- backward_step_wrapper(obs_df, mt)
```

## Use Richardson-Lucy deconvolution to infer epicurve from synthetic observations
```{r}
source('../code/Richardson_Lucy.R')
get_delay_multinom <- function(r_delay){
  data_frame(delay = r_delay(10000) %>% round) %>%
  group_by(delay) %>%
  summarise(n = n()) %>%
  ungroup() %>%
  complete(delay = 0:max(delay), fill = list(n = 0)) %>%
  mutate(n = n/sum(n)) %>%
  pull(n)
}

## Add deconvolution of observed cases
inf_obs_df <- inf_obs_df %>%
  merge(get_RL(observed = inf_obs_df$obs_case, 
          times = inf_obs_df$time, 
          p_delay = get_delay_multinom(case_delay_dist), 
          out_col_name = 'case_deconvolved'),
by = 'time') %>%
## Add deconvolution of deaths
    merge(get_RL(observed = inf_obs_df$obs_deaths, 
          times = inf_obs_df$time, 
          p_delay = get_delay_multinom(death_delay_dist), 
          out_col_name = 'death_deconvolved'),
by = 'time')
```

# Fig. 3A

* Plot true infections (from synthetic data)
* Plot true observations (from forward step)
* Plot back-calculated infection curves using the shift or sample method
* For back-calculated infection curves using sample method, compare resuts with or without adjustment for right censoring.
```{r}
inf_obs_df %>%
  select(time, incidence, contains('case')) %>%
  pivot_longer(-time) %>%
  filter(name != 'sample_case') %>%
  mutate(name = factor(name,
                       levels = c('incidence', 'obs_case', 'sample_case_unadjusted', 'case_deconvolved', 'shift_case'),
                       labels = c('infections', 'observed cases', 'convolve', 'deconvolve', 'shift'))) %>%
  ggplot() +
  #geom_line(data = inf_obs_df, aes(x = time, y = sample_case_unadjusted), color = 'mediumorchid3', lty = 2)+
  geom_line(aes(x = time, y = value, color = name)) +
  xlim(c(40, mt))+
  #ylim(c(750, 7000))+
  scale_color_manual(values = c('orange', 'mediumseagreen', 'mediumorchid3', 'red4', 'dodgerblue'))+
  theme(legend.position = 'none')+
  ylab('count')+
  labs(color = '') -> cases

cases
```


# Fig. 3B - deaths
```{r}
inf_obs_df %>%
  select(time, incidence, contains('death')) %>%
  pivot_longer(-time) %>%
  filter(name != 'sample_death') %>%
  mutate(name = factor(name,
                       levels = c('incidence', 'obs_deaths', 'shift_death', 'sample_death_unadjusted', 'death_deconvolved'),
                       labels = c('infections', 'observed', 'shift', 'convolve', 'deconvolve'))) %>%
  ggplot() +
  #geom_line(data = inf_obs_df, aes(x = time, y = sample_death_unadjusted), color = 'mediumorchid3', lty = 2)+
  geom_line(aes(x = time, y = value, color = name)) +
  xlim(c(40, mt))+
  #ylim(c(750, 7000))+
  scale_color_manual(values = c('orange', 'mediumseagreen', 'dodgerblue', 'mediumorchid3', 'red4'))+
  theme(legend.position = 'none')+
  labs(color = '') -> deaths
deaths
```




# Fig. 3C Assess how errors in inferred times of infection impact $R_t$

## Wrappers to estimate Rt using cases observed up to a given time point

```{r}
get_Rts <- function(df.in, wdw){
  rt_df <- df.in %>%
    merge(get_cori(df.in, icol_name = 'incidence', window = wdw, out_name = 'incidence'), by = 'time', all.x = TRUE) %>%  ## Add Rt fit to true infection incidence
    merge(get_cori(df.in, icol_name = 'shift_case', window = wdw, out_name = 'shift_case'), by = 'time', all.x = TRUE) %>%
    merge(get_cori(df.in, icol_name = 'shift_death', window = wdw, out_name = 'shift_death'), by = 'time', all.x = TRUE)
}

Rt_give_data_to_time_T <- function(df.in, max_time, window){
  df1 <- backward_step_wrapper(df.in, mt)
  get_Rts(df1, wdw = window)
}
```

## Estimate Rt using cases up to time 100
```{r}
Rt_to_T <- Rt_give_data_to_time_T(df.in = obs_df, max_time = mt, window = 1) %>% as.tbl()
```

```{r}
ymax <- 2.5
Rt_to_T %>%
  pivot_longer(cols = contains('.'), names_to = c('strategy', '.value'), names_sep = '\\.') %>%
  mutate(strategy = factor(strategy, levels = c('incidence', 'shift_case', 'shift_death'))) %>%
  group_by(strategy) %>%
  mutate(flag = ifelse(mean>0 & !is.na(mean) & is.finite(mean), TRUE, FALSE),
         maxT = max(time[flag]),
         laglab = sprintf('%.0fd gap', ceiling(as.numeric(max(time)-maxT)))) %>%
  mutate(`025` = ifelse(`025`>ymax, ymax, `025`),
         `975` = ifelse(`975`>ymax, ymax, `975`)) %>%
  filter(strategy != 'incidence') %>%
  ggplot(aes(x = time)) +
  geom_line(aes(y = true_rt), lwd = 1) +
  geom_line(aes(y = mean, color = strategy), lwd = .75) +
  geom_ribbon(aes(ymin = `025`, ymax = `975`, fill = strategy), alpha = 0.3) +
  geom_vline(aes(xintercept = mt), lty = 2) +
  geom_vline(aes(xintercept = maxT, color = strategy), lty = 2) +
  ylab(expression(paste(R[t])))+
  xlab('day')+
  ylim(c(.5,ymax))+
  xlim(c(50, mt))+
  #ggtitle('Accuracy of Rt estimates by data type',
  #        subtitle = 'Estimated using "shift" method in panel B') +
  theme(legend.position = 'none')+
  labs(color = 'estimated from shifted', fill = 'estimated from shifted') -> a
a
```


## Save individual panels
```{r}
cowplot::plot_grid(
  get_legend(
  deaths + theme(legend.box.margin = margin(6, 0, 0, 6),
          legend.position = 'bottom')
)
)
ggsave('figs/backcalc_legend.png', width = 7, height = 1.5)

ggsave(plot = cases, 
       sprintf('figs/cases.png'), 
       width = 3.5, 
       height = 3.5, 
       units = 'in', 
       dpi = 300)

ggsave(plot = deaths, 
       sprintf('figs/deaths.png'), 
       width = 3.5, 
       height = 3.5, 
       units = 'in', 
       dpi = 300)


ggsave(plot = a, 
       'figs/Rt_ests.png', 
       width = 7, 
       height = 4.5, 
       units = 'in', 
       dpi = 300)
```


## For presentations
```{r}
inf_obs_df %>%
  select(time, incidence, contains('case')) %>%
  pivot_longer(-time) %>%
  filter(name != 'sample_case') %>%
  mutate(name = factor(name,
                       levels = c('incidence', 'obs_case', 'shift_case', 'sample_case_unadjusted'),
                       labels = c('infections', 'observed cases', 'shift', 'convolve'))) %>%
  ggplot() +
  #geom_line(data = inf_obs_df, aes(x = time, y = sample_case_unadjusted), color = 'mediumorchid3', lty = 2)+
  geom_line(aes(x = time, y = value, color = name)) +
  xlim(c(40, mt))+
  #ylim(c(750, 7000))+
  scale_color_manual(values = c('orange', 'mediumseagreen', NA, NA))+
  theme(legend.position = 'none')+
  ylab('count')+
  labs(color = '') -> cases1

ggsave(plot = cases1, 
       sprintf('figs/cases0.png'), 
       width = 3.5, 
       height = 3.5, 
       units = 'in', 
       dpi = 300)


inf_obs_df %>%
  select(time, incidence, contains('case')) %>%
  pivot_longer(-time) %>%
  filter(name != 'sample_case') %>%
  mutate(name = factor(name,
                       levels = c('incidence', 'obs_case', 'shift_case', 'sample_case_unadjusted'),
                       labels = c('infections', 'observed cases', 'shift', 'convolve'))) %>%
  ggplot() +
  #geom_line(data = inf_obs_df, aes(x = time, y = sample_case_unadjusted), color = 'mediumorchid3', lty = 2)+
  geom_line(aes(x = time, y = value, color = name)) +
  xlim(c(40, mt))+
  #ylim(c(750, 7000))+
  scale_color_manual(values = c('orange', 'mediumseagreen', 'dodgerblue', NA))+
  theme(legend.position = 'none')+
  ylab('count')+
  labs(color = '') -> cases2

ggsave(plot = cases2, 
       sprintf('figs/cases01.png'), 
       width = 3.5, 
       height = 3.5, 
       units = 'in', 
       dpi = 300)


inf_obs_df %>%
  select(time, incidence, contains('death')) %>%
  pivot_longer(-time) %>%
  filter(name != 'sample_death') %>%
  mutate(name = factor(name,
                       levels = c('incidence', 'obs_deaths', 'shift_death', 'sample_death_unadjusted'),
                       labels = c('infections', 'observed', 'shift', 'convolve'))) %>%
  ggplot() +
  #geom_line(data = inf_obs_df, aes(x = time, y = sample_death_unadjusted), color = 'mediumorchid3', lty = 2)+
  geom_line(aes(x = time, y = value, color = name)) +
  xlim(c(40, mt))+
  #ylim(c(750, 7000))+
  scale_color_manual(values = c('orange', 'mediumseagreen', NA, NA))+
  theme(legend.position = 'none')+
  labs(color = '') -> deaths1

ggsave(plot = deaths1, 
       sprintf('figs/deaths0.png'), 
       width = 3.5, 
       height = 3.5, 
       units = 'in', 
       dpi = 300)



inf_obs_df %>%
  select(time, incidence, contains('death')) %>%
  pivot_longer(-time) %>%
  filter(name != 'sample_death') %>%
  mutate(name = factor(name,
                       levels = c('incidence', 'obs_deaths', 'shift_death', 'sample_death_unadjusted'),
                       labels = c('infections', 'observed', 'shift', 'convolve'))) %>%
  ggplot() +
  #geom_line(data = inf_obs_df, aes(x = time, y = sample_death_unadjusted), color = 'mediumorchid3', lty = 2)+
  geom_line(aes(x = time, y = value, color = name)) +
  xlim(c(40, mt))+
  #ylim(c(750, 7000))+
  scale_color_manual(values = c('orange', 'mediumseagreen', 'dodgerblue', NA))+
  theme(legend.position = 'none')+
  labs(color = '') -> deaths2

ggsave(plot = deaths2, 
       sprintf('figs/deaths01.png'), 
       width = 3.5, 
       height = 3.5, 
       units = 'in', 
       dpi = 300)


```


[1]: https://www.mdpi.com/2077-0383/9/2/538
